Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-03-20 arcticgrids.tex
updates

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington \affil{1}, Marcus Lofverstrom \affil{2}, Peter H. Lauritzen \affil{1} and Andrew Gettelman \affil{1}}

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}, which refers to numerical instability arising from the convergence of meridians to polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study, a spectrum of model grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}) are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the meteorology over the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole-problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems unlikely (describe why) and is explored further in this study.

An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c). This grid flexibility also allows for variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}). Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space). However, unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and melting) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes. Moreover, GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

@@

The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a number of different atmospheric dynamical cores. These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores in the current model. FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study. As such, the results presented in this study are comparing the performance of the SE and FV dycores.

@@

The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,DetAl2012IJHPCA}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}). Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points. This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary. Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.

@@

The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids. (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by about an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so  the NCAR topography software \cite{gmdd-8-4623-2015} was modified to scale the smoothing radius by the local element size.

@@

Six grid are evaluated in this study (Table X). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

@@

@@

The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used for all simulations in this study. CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

The CAM convention is that the physics package determines the vertical resolution. Since all runs use CAM6 physics, all grids and dycores use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. The physics time-step, in contrast, is dependent on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step should be reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids.

Initial simulations with the $ne30pg3$ spectral-element grid produced a weaker shortwave cloud forcing relative to the tuned up finite-volume dycore. All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic shortwave cloud forcing. These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6. Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy. For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.

@@

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2. The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}. For land ice tiles coinciding with the Antarctic and Greenland Ice Sheets, the surface energy and mass balance are evaluated at ten elevation classes, and integrated over sub-grid hypsometric area-elevation bins for more accurate estimation of the surface mass balance of the big ice sheets. Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation. Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years. The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

@@

The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface. All these processes are represented by different components in CESM, and the surface mass balance is computed by CLM. CLM uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation model. The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, allowing for the precipitation rates to be repartitioned into solid or liquid based on the temperature of each EC. This technique allows for a unique surface energy balance and surface mass balance at each EC. Integrating over all ECs using the area weights then provides a more accurate SMB. For a detailed description of how the SMB is computed in the CESM, the reader is referred to \cite{SETAL2019TC,KETAL2020JAMES}.

@@

Simulated SMB are compared against several observational data products. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations that have been forced by reanalysis at its domain boundaries. The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}. The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset. The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.

A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset. While Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, we seek to compare these integrated quantities with RACMO. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study. In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

@@

Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed. Figure~\ref{fig:dT-lores} shows plots of the 1979-1998 annual mean, zonal mean height expressed as differences between the uniform resolution grids and dycores. The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB. The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.

@@

As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}). The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes. The differences in temperature between $ne30pg3$ and $ne30pg2$ are small, with a slight warming near the tropopause at high latitudes. This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.

@@

>>>>

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.

An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c). This grid flexibility also permits variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}). Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space). Unstructured grids also scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

@@

The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports a number of different atmospheric dynamical cores. These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores. FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study. As such, the results presented in this study are comparing the performance of the SE and FV dycores.

@@

The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}). Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points. This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary. Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.

@@

The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids. (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.

@@

Six grid are evaluated in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

@@

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ARCTIC$ & SE & 111 & 28 & 450 \\
   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step. The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
 \label{tbl:table1}
 \end{table*}

@@

The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study. CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. The physics time-step is dependent on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.

Initial simulations with the $ne30pg3$ spectral-element grid produced weaker shortwave cloud forcing relative to the tuned up finite-volume dycore. All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic cloud forcing and top-of-atmosphere radiation balance. These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6. Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy. For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.

@@

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface. These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset. The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC. Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice. A unique surface energy balance and SMB is computed for each EC. Integrating over all ECs using the area weights provides a more accurate SMB. For a more detailed description of how the SMB is computed in CESM, the reader is referred to \cite{CISM1,SETAL2019TC,KETAL2020JAMES}.

Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over a 20 year $ARCTIC$ $FHIST$ run for about 500 years. The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

\subsection{Observational Datasets}

Several observational datasets are used in this study to understand the performance of the simulations. A list of the datasets used in this study are shown in Table~\ref{tbl:table2}. Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets are gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalysis at its domain boundaries. The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}. The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset. The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccc}
   \hline
   data product & years used in this study & citation \\ 
   \hline
   ERA5 & 1979-1998 & \cite{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & \cite{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & \cite{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & \cite{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & \cite{NETAL2019SCIENCE} \\
   LIVVkit,v2 & 1979-1998 & \cite{LIVVkit} \\
 \hline
 \end{tabular}
  \caption{Description of observational datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

@@

A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.

Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, but we seek to compare these integrated quantities to RACMO2.3. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study. In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

@@

@@

Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores. The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB. The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.

@@

As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}). The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes. Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes. This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.

@@

\subsection{Shortwave radiation over Greenland}

In addition to summer temperatures, shortwave radiation is also an important determinant of snow/ice melt. Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface, zoomed in over Greenland. The top panel computes the bias using the CERES-EBAF dataset, and the bottom panel using RACMO2.3p2 dataset.
This halo of excessive incident shortwave radiation around the coasts of Greenland is apparent in both datasets, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.

The interior of the ice sheet receives too little shortwave radiation in the coarser grids. In the variable-resolution grids, both the interior deficit in shortwave and the excessive shortwave around the oceanic perimeter of Greenland are improved. This suggests that the coarse grids clouds are too thick in the interior of Greenland, and too thin around the perimeter of Greenland, and that increasing horizontal resolution balances out these biases. This is consistent with total summer cloud fraction bias, computed from the CALIPSO cloud dataset (Figure~\ref{fig:prect}). Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates any changes arising from any single layer due to the maximum overlap assumption used to compute this quantity. Despite the attenuated signal, the total cloud fraction does indicate a reduction in cloud coverage in the interior, and an increase in cloudiness about the oceanic perimeter, in the variable-resolution grids.
